---
title: 'Reproducible Research Project #1'
author: "Andrew Holshouser"
date: "May 31, 2017"
output: html_document
---

##Loading the data

```{r loading}
wlk<-read.csv('activity.csv')
```

##Histogram of the total number of steps taken each day

```{r hist1}
wlkDay<-aggregate(steps~date,wlk, sum, na.rm=T)
hist(wlkDay$steps, main='Total # of Steps Per Day', xlab='# of Steps')
```

##Mean and median number of steps taken each day
```{r MeanMedian}
wlkMean<-mean(wlkDay$steps)
wlkMedian<-median(wlkDay$steps)
```

Result:
```{r MeanMedian Result, echo=FALSE}
print(wlkMean) 
print(wlkMedian)
```

##Time series plot of the average number of steps taken
```{r TimeSeries Plot}
wlkTS<-aggregate(steps~interval, wlk, mean, na.rm=T)
plot(wlkTS, main='Time Series by Avg. number of steps', pch=16)
lines(wlkTS)
```

##The 5-minute interval that, on average, contains the maximum number of steps
```{r MaxInterval}
wlkTSMax<-wlkTS[which.max(wlkTS$steps),]
```

Result:
```{r MaxIntervalResult, echo=FALSE}
print(wlkTSMax)
```

##Code to describe and show a strategy for imputing missing data

Identify the NA values within the data set
```{r IdentifyNA}
wlkNA<-is.na(wlk$steps) ##Identify missing values (NA)
wlktotNA<-sum(wlkNA) ##Count the NA values
```

Result: # of values with NA
```{r IdentifyNAResult, echo=FALSE}
print(wlktotNA)
```

Replace missing interval with the mean of the 5-min interval
```{r ReplaceNA}
wlk0<-read.csv('activity.csv')
wlk0$steps[is.na(wlk0$steps)]=wlkTS$steps
```

Result: # of value with NA after replacement
```{r ReplaceTot, echo=FALSE}
sum(is.na(wlk0$steps))
```

##Histogram of the total number of steps taken each day after missing values are imputed
```{r Hist2}
wlk0Day<-aggregate(steps~date,wlk0, sum)
hist(wlk0Day$steps, main='Total # of Steps Per Day (Non-NA)', xlab='# of Steps')
```

##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r CompareWEWD}
library(chron)
wlk$we=chron::is.weekend(wlk$date) ##TRUE is a weekend / FALSE is a weekday
wlkWE<-subset(wlk, wlk$we==TRUE)
wlkWD<-subset(wlk, wlk$we==FALSE)

wlkWEDay<-aggregate(steps~date, wlkWE,mean)
wlkWDDay<-aggregate(steps~date, wlkWD,mean)

par(mfrow=c(1,2), mar=c(4,4,2,1))
hist(wlkWDDay$steps, main='Weekday', xlab='Avg. # of Steps',col="blue", 
     ylim=c(0,12))
hist(wlkWEDay$steps, main='Weekend Day ', xlab='Avg. # of Steps',col="red", 
     ylim=c(0,12))
```
